---
title: "Data Wrangling"
author: "Emma Kaufman and Jaime Wargo"
date: "2024-03-26"
output: pdf_document
---
```{r package, message=FALSE, warning=FALSE}
library(lubridate)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(outliers)
library(tidyverse)
library(smooth)
library(zoo)
library(kableExtra)
library(readxl)
library(here)
```

```{r}
load <- read_excel("Data_TOPOST/load.xlsx")
relative_humidity <- read_excel("Data_TOPOST/relative_humidity.xlsx")
temperature <- read_excel("Data_TOPOST/temperature.xlsx")
```

## WRANGLE/PROCESS THE DATASET

You will need to transform hourly data into daily data. See the Rmd file from Lesson 11 for instruction on how to aggregate your dataset using pipes. You should take the **average** of the 24 hours to obtain the daily averages.
```{r}
#start with wrangling load data
load_all <- load %>% 
  pivot_longer(cols = h1:h24) %>% 
  rename(Hour = name,
         Load= value) %>% 
  mutate(Hour= as.numeric(gsub('h','', Hour))-1) %>% 
  mutate(date= ymd(date)) %>% 
  mutate(Day = day(date),
         Month= month(date),
         Year= year(date)) %>% 
  select(meter_id, date, Year, Month, Day, Hour, Load)

#check for NAs
summary(load_all$Load)

```

```{r}
#Creating a data frame with daily observations
load_daily <- load_all %>% 
  filter( !is.na(Load) ) %>% 
  group_by(meter_id, date, Year, Month, Day) %>% # here we left column with hour out to calculate daily mean
  summarise( Daily_mean_load = mean(Load)) #take the mean for the day

ggplot(load_daily, aes(x=date,y=Daily_mean_load)) +
  geom_line() +
  ylab("Average Daily Load")

#check for NAs
summary(load_daily$Daily_mean_load)
```
```{r}
#temperature and relative humidity data
temperature_all <- temperature %>% 
  mutate(date= ymd(date)) %>% 
  mutate(Day= day(date),
          Month= month(date),
          Year= year(date)) %>% 
  mutate(hr= hr-1) %>% 
  rename(hour= hr) 

relative_humidity_all <- relative_humidity %>% 
   mutate(date= ymd(date)) %>% 
  mutate(Day= day(date),
          Month= month(date),
          Year= year(date)) %>% 
  mutate(hr= hr-1) %>% 
  rename(hour= hr)

#check for NA
summary(temperature_all)
summary(relative_humidity_all)

#summarize daily data
temp_daily <- temperature_all %>% 
  group_by(date, Year, Month, Day) %>% # here we left column with hour out to calculate daily mean
  #daily mean for each station
  summarise_at(vars(matches("^t_ws[1-9]|^t_ws1[0-8]$")), list(Daily_temp = ~mean(., na.rm = TRUE)))
 

rh_daily <- relative_humidity_all %>% 
  group_by(date, Year, Month, Day) %>% # here we left column with hour out to calculate daily mean
  #daily mean for each station
  summarise_at(vars(matches("^rh_ws[1-9]|^rh_ws1[0-8]$")), list(Daily_temp = ~mean(., na.rm = TRUE)))

```

```{r}
#save datasets

write.csv(load_daily,file=here("Data","Processed","Daily_Load.csv"))
write.csv(temp_daily,file=here("Data","Processed","Daily_Temp.csv"))
write.csv(rh_daily,file=here("Data","Processed","Daily_Relative_Humidity.csv"))
```

## CREATE A TIME SERIES OBJECT

After you process your dataset use the `msts()` function to create a time series object. You need to use `msts()` instead of `ts()` because your daily data will have more than one seasonal component. After you process your dataset use the `msts()` function to create a time series object.

```{r}
#load time series object
ts_load_daily <- msts(load_daily$Daily_mean_load, #daily data frame mean
                           seasonal.periods =c(7,365.25), #seasonal periods
                           start=c(2005,1,1))
```

```{r}
#decomposition of daily data
ts_load_daily %>% mstl() %>%
  autoplot()
```

